A Large and Variable Leading Tail of Helium in a Hot Saturn Undergoing Runaway Inflation

### 2023-09-07 paper/paper1/ms.tex
orbital evolution

\edit1{\subsection{Prospect for orbital evolution}
  Evaporating planets may lose a large fraction of their mass through Roche lobe overflow \citep{2014ApJ...793L...3V,2015ApJ...813..101V}.  These profoundly eroded planets should exhibit significant orbital evolution.  The extent and direction of such orbital evolution depends on the specific angular momentum loss, integrated over the mass loss history of the planet.  The observations presented here indicate a mass loss vector perpendicular to the orbit and towards the star.  We can reconstruct the starting orbit if we assume the current mass loss vector has been constant throughout its mass loss history.  Following \citet{2014ApJ...793L...3V} and \citet{2023ApJ...945L..36T} we can constrain the change in semi-major axis, a:
  \begin{equation}
    \frac{\dot{a}}{a} = -2 \frac{\dot{M}}{M}\left (1 - \frac{M}{M_\star} \right ).
  \end{equation}
}

>>>>

\edit1{\subsection{Prospect for orbital evolution}
  Evaporating planets may lose a large fraction of their mass through Roche lobe overflow \citep{2014ApJ...793L...3V,2015ApJ...813..101V}.  These profoundly eroded planets should exhibit significant orbital evolution.  The extent and direction of such orbital evolution depends on the specific angular momentum loss, integrated over the mass loss history of the planet.  The observations presented here indicate a mass loss vector perpendicular to the orbit and towards the star.  A large enough mass loss rate would be evinced as a changing orbital period $\dot{P}$, which would manifest as a weak linear trend in Transit Timing Variations (TTVs).  \citet{2017AJ....153..211Z} searched for TTVs in their ground based photometry, which spanned a baseline of about 160 days with 6 data points.  The \emph{TESS} photometry provides over 24 full or partial transits over 8 years later, yielding a much larger lever arm for gauging the prospect of TTVs.  Table \ref{tabOrbit} compares the period determined separately between these two datasets.  The TESS-based period is 1-$\sigma$ longer than the \citet{2017AJ....153..211Z} period.  A joint orbit fit of both the original lightcurves, existing and planned TESS lightcurves, and higher precision RVs will be able to better constrain orbital evolution.  A non-detection of TTVs could still provide meaningful constraints on extreme mass loss scenarios.
}
